---
title: 'Marine Aquaculture'
author: "Melissa Widas"
date: "2022-11-16"
output:
    html_document:
      print_df: paged
      toc: yes
      toc_depth: 4
      toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
getwd()
```

## Overview
Marine aquaculture has the potential to play an important role in the global food supply as a more sustainable protein option than land-based meat production.[^hall] [Gentry et al.](https://www.nature.com/articles/s41559-017-0257-9) mapped the potential for marine aquaculture globally based on multiple constraints, including ship traffic, dissolved oxygen, bottom depth .[^gentry]

We will be determining which Exclusive Economic Zones (EEZ) on the West Coast of the US are best suited to developing marine aquaculture for several species of oysters.\

Based on previous research, we know that oysters needs the following conditions for optimal growth:\

-   sea surface temperature: 11-30&deg;C\
-   depth: 0-70 meters below sea level

##### Learning objectives:
-   combining vector/raster data\
-   resampling raster data\
-   masking raster data\
-   map algebra\

### Data
#### Sea Surface Temperature
We will use average annual sea surface temperature (SST) from the years 2008 to 2012 to characterize the average sea surface temperature within the region. The data we are working with was originally generated from [NOAA's 5km Daily Global Satellite Sea Surface Temperature Anomaly v3.1](https://coralreefwatch.noaa.gov/product/5km/index_5km_ssta.php).

#### Bathymetry
To characterize the depth of the ocean we will use the [General Bathymetric Chart of the Oceans (GEBCO)](https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area).[^gebco]

#### Exclusive Economic Zones
We will be designating maritime boundaries using Exclusive Economic Zones off of the west coast of US from [Marineregions.org](https://www.marineregions.org/eez.php).

## Assignment

Below is an outline of the steps you should consider taking to achieve the assignment tasks. 

#### Prepare data (5 points)
To start, we need to load all necessary data and make sure it has the coordinate reference system. 

-   load necessary packages and set path\ 
    -   I recommend using the [`here` package](https://here.r-lib.org/)
-   read in the shapefile for the West Coast EEZ (`wc_regions_clean.shp`)\
-   read in SST rasters
    -   `average_annual_sst_2008.tif`\
    -   `average_annual_sst_2009.tif`\
    -   `average_annual_sst_2010.tif`\
    -   `average_annual_sst_2011.tif`\
    -   `average_annual_sst_2012.tif`\
-   combine SST rasters into a raster stack\
-   read in bathymetry raster (`depth.tif`)\
-   check that data are in the same coordinate reference system\
    -   reproject any data not in the same projection\

```{r include=TRUE, warning=FALSE, message=FALSE}
# load packages
library(here)
library(tidyverse)
library(sf)
library(terra)
library(ggspatial)
library(patchwork)

# read in the shapefile for the West Coast EEZ (`wc_regions_clean.shp`)
wc <- read_sf(here("data","wc_regions_clean.shp"))

# read in SST rasters
sst_2008 <- rast(here("data", "average_annual_sst_2008.tif"))
sst_2009 <- rast(here("data", "average_annual_sst_2009.tif"))
sst_2010 <- rast(here("data", "average_annual_sst_2010.tif"))
sst_2011 <- rast(here("data", "average_annual_sst_2011.tif"))
sst_2012 <- rast(here("data", "average_annual_sst_2012.tif"))

# read in depth raster
depth <- rast(here("data", "depth.tif"))
```

#### Process data (10 points)
Next, we need process the SST and depth data so that they can be combined. In this case the SST and depth data have slightly different resolutions, extents, and positions. We don't want to change the underlying depth data, so we will need to resample to match the SST data using the nearest neighbor approach. 

-   find the mean SST from 2008-2012\
-   convert SST data from Kelvin to Celsius\
    -   hint: subtract by 273.15\
-   crop depth raster to match the extent of the SST raster\
-   note: the resolutions of the SST and depth data do not match\
    -   resample the NPP data to match the resolution of the SST data using the nearest neighbor approach\
-   check that the depth and SST match in resolution, extent, and coordinate reference system\
    -   hint: can the rasters be stacked?
    
```{r include=TRUE}
# stack annual data 
sst <- c(sst_2008, sst_2009, sst_2010, sst_2011, sst_2012)


# reproject Coordinate System
st_crs(sst)     # WGS84, EPSG 9122
st_crs(wc)       # WGS84 EPSG 4326
st_crs(depth)         # WGS84, EPSG 4326

sst <- project(sst, depth)
st_crs(sst)     # WGS84, EPSG 4326


# Find mean Sea Surface Temperature

# Collapse down layers to one SST mean and convert to celcius
sst_mean <- mean(sst)
sst_mean <- (sst_mean$mean - 273.15)

# Check they will match
sst_depth <- c(depth, sst_mean)

depth_crop <- crop(depth, sst_mean)

depth_crop_res <- resample(x= depth_crop, y = sst_mean, method = "near")

# stack annual data 
sst_depth <- c(sst_mean, depth_crop_res)
sst_depth_rast <- rast(sst_depth)
```

#### Find suitable locations (20)
In order to find suitable locations for marine aquaculture, we'll need to find locations that are suitable in terms of both SST and depth.

-   reclassify SST and depth data into locations that are suitable for Lump sucker fish\
    -   hint: set suitable values to `1` and unsuitable values to `NA`\
-   find locations that satisfy both SST and depth conditions\
    -   hint: create an overlay using the `lapp()` function multiplying cell values\
    
```{r include=TRUE}
# Reclassification for sst 
sst_reclass <- matrix(c(-Inf, 11, NA,
                   11, 30, 1,        # sea surface temperature: 11-30°C
                   30, Inf, NA),
                     ncol = 3, byrow = TRUE)
oyster_sst <- classify(sst_mean, rcl = sst_reclass)
plot(oyster_sst, col = "pink", bty = "L",
     main = "Sea Surface Temperature, Oysters")
```

```{r}
# Reclassification for depth
depth_reclass <- matrix(c(-Inf, -70, NA,
                     -70, 0, 1,      # depth: 0-70 meters below sea level
                     0, Inf, NA),
                     ncol = 3, byrow = TRUE)
oyster_depth <- classify(depth_crop_res, rcl = depth_reclass)
plot(oyster_depth, col = "pink", bty = "L",
     main = "Depth for Oysters")
```

```{r}
# Overlay to find areas that satisfy BOTH 
fun_mult = function(x,y){return(x*y)}     # function to multiply layers

oyster_both <- lapp(c(oyster_sst, oyster_depth), fun_mult)
plot(oyster_both, col = "pink", bty = "L",
     main = "Habitat for Oysters")
```

#### Determine the most suitable EEZ (20 points)
We want to determine the total suitable area within each EEZ in order to rank zones by priority. To do so, we need to find the total area of suitable locations within each EEZ.

-   select suitable cells within West Coast EEZs\
-   find area of grid cells\
-   find the total suitable area within each EEZ\
    -   hint: it might be helpful to rasterize the EEZ data\
-   find the percentage of each zone that is suitable\
    -   hint it might be helpful to join the suitable area by region onto the EEZ vector data\
    
```{r include=TRUE}
# Turn Regions data into raster 
wc$rgn <- as.factor(wc$rgn)
wc_raster <- rasterize(vect(wc), oyster_both, field = "rgn") # use vect to create a spatial vector


# Select suitable cells in EEZ 
suitable_mask <- mask(wc_raster, oyster_both)


# Find total suitable area within each EEZ 
oyster_area <- expanse(suitable_mask, unit = "km", byValue = TRUE) # use expanse to calculate the area


# Find percentage of each zone
oyster_region <- cbind(oyster_area, wc)
oyster_region <- oyster_region %>% 
  mutate(percent = (area / area_km2)*100)
```
    
#### Visualize results (5 points)
Now that we have results, we need to present them!

Create the following maps:

-   total suitable area by region\
-   percent suitable area by region\

Include:

-   legible legends\
-   updated color aesthetics\
-   basemap\

```{r include=TRUE, warning=FALSE}
# create visualization for total suitable area by region
total_area_region <- ggplot() +
  geom_sf(data = oyster_region, # plot the oyster regions
          aes(fill = area, geometry = geometry)) + # fill by area
  geom_sf_text(data = oyster_region, aes(geometry = geometry, label = value)) +
 labs(fill='Area (m)') + # alter legend title
  scale_fill_viridis_c(begin = 0.4) +
  theme_void() +
  annotation_north_arrow( # add north arrow
  height = unit(1.5, "cm"),
  width = unit(1.5, "cm"),
  pad_x = unit(1, "cm"),
  pad_y = unit(1, "cm"),
  rotation = NULL,
  style = north_arrow_fancy_orienteering
) +
  annotation_scale( # add scale bar
  plot_unit = NULL,
  bar_cols = c("black", "white"),
  line_width = 1,
  height = unit(0.15, "cm"),
  pad_x = unit(2, "cm"),
  pad_y = unit(0.15, "cm"),
  text_pad = unit(0.15, "cm"),
  text_cex = 0.7,
  text_face = NULL,
  text_family = "",
  tick_height = 0.6
) +
  labs(title= "Total Suitable Area for Oysters", color = "") # add map title 
  
total_area_region
```
```{r, warning=FALSE}
# create visualization for percent suitable area by region
percent_region <- ggplot() +
  geom_sf(data = oyster_region, # plot the oyster regions
          aes(fill = percent, geometry = geometry)) + # fill by percent
  geom_sf_text(data = oyster_region, aes(geometry = geometry, label = value)) +
  scale_fill_viridis_c(begin = 0.4) +
 labs(fill='Percentage') +
  theme_void() +
  annotation_north_arrow( # add north arrow
  height = unit(1.5, "cm"),
  width = unit(1.5, "cm"),
  pad_x = unit(1, "cm"),
  pad_y = unit(1, "cm"),
  rotation = NULL,
  style = north_arrow_fancy_orienteering
) +
  annotation_scale( # add scale bar
  plot_unit = NULL,
  bar_cols = c("black", "white"),
  line_width = 1,
  height = unit(0.15, "cm"),
  pad_x = unit(2, "cm"),
  pad_y = unit(0.15, "cm"),
  text_pad = unit(0.15, "cm"),
  text_cex = 0.7,
  text_face = NULL,
  text_family = "",
  tick_height = 0.6
) +
  labs(title= "Suitable Percentage for Oysters", color = "") # add map title
  
percent_region
```

```{r, warning=FALSE}
total_area_region+percent_region
```

#### Broaden your workflow! (40 points)
Now that you've worked through the solution for one group of species, let's update your workflow to work for other species. Please create a function that would allow you to reproduce your results for other species. Your function should be able to do the following:\

-   accept temperature and depth ranges and species name as inputs\
-   create maps of total suitable area and percent suitable area per EEZ with the species name in the title\

Run your function for a species of your choice! You can find information on species depth and temperature requirements on [SeaLifeBase](https://www.sealifebase.ca/search.php). Remember, we are thinking about the potential for marine aquaculture, so these species should have some reasonable potential for commercial consumption. 

Chose to run the function on a species of shrimp, commonly known as pink shrimp, Pandalus jordani. Pink Shrimp are found at these conditions from [NOAA](https://www.fisheries.noaa.gov/species/pink-shrimp):

-   sea surface temperature: 6-12&deg;C\
-   depth: 36-437 meters below sea level

```{r, warning=FALSE}
aquaculture_regions <- function(species_name, min_temp, max_temp, max_depth, min_depth){
  
  # reclassify
  sst_reclass <- matrix(c(-Inf, min_temp, NA,
                          min_temp, max_temp, 1,
                          max_temp, Inf, NA),
                        ncol = 3, byrow = TRUE)
  depth_reclass <- matrix(c(-Inf, max_depth, NA,
                            max_depth, min_depth, 1,
                            min_depth, Inf, NA),
                          ncol = 3, byrow = TRUE)  
  
  # apply reclassify
  species_sst <- classify(sst_mean, rcl = sst_reclass)
  species_depth <- classify(depth, rcl = depth_reclass)
  fun_mult = function(x,y){return(x*y)}     # function to multiply layers
  species_suitable <- lapp(c(species_sst, species_depth), fun_mult)
  
  # Turn Regions data into raster 
  wc$rgn <- as.factor(wc$rgn)
  wc_raster <- rasterize(vect(wc), species_suitable, field = "rgn")
  
  
  # Select suitable cells in EEZ 
  species_mask <- mask(wc_raster, species_suitable)
  
  
  # Find total suitable area within each EEZ 
  species_area <- expanse(species_mask, unit = "km", byValue = TRUE)
  
  
  # Find percentage of each zone
  species_region <- cbind(species_area, wc)
  species_region <- species_region %>%  
    mutate(percent = (area / area_km2)*100)
  
  # create visualization for total suitable area by region
  total_area_region <- ggplot() +
    geom_sf(data = species_region, # plot the oyster regions
            aes(fill = area, geometry = geometry)) + # fill by area
    geom_sf_text(data = species_region, aes(geometry = geometry, label = value)) +
    scale_fill_viridis_c(begin = 0.4) +
    labs(fill='Area (m)') + # alter legend title
    theme_void() +
    annotation_north_arrow( # add north arrow
      height = unit(1.5, "cm"),
      width = unit(1.5, "cm"),
      pad_x = unit(1, "cm"),
      pad_y = unit(1, "cm"),
      rotation = NULL,
      style = north_arrow_fancy_orienteering
    ) +
    annotation_scale( # add scale bar
      plot_unit = NULL,
      bar_cols = c("black", "white"),
      line_width = 1,
      height = unit(0.15, "cm"),
      pad_x = unit(2, "cm"),
      pad_y = unit(0.15, "cm"),
      text_pad = unit(0.15, "cm"),
      text_cex = 0.7,
      text_face = NULL,
      text_family = "",
      tick_height = 0.6
    ) +
    labs(title = print(paste0("Total Suitable Area for ", species_name))) # add map title 
         
  total_area_region
         
  # create visualization for total suitable area by region
  percent_region <- ggplot() +
    geom_sf(data = species_region, # plot the oyster regions
            aes(fill = percent, geometry = geometry)) + # fill by area
    geom_sf_text(data = species_region, aes(geometry = geometry, label = value)) +
    scale_fill_viridis_c(begin = 0.4) +
    labs(fill='Percentage') + # alter legend title
    theme_void() +
    annotation_north_arrow( # add north arrow
      height = unit(1.5, "cm"),
      width = unit(1.5, "cm"),
      pad_x = unit(1, "cm"),
      pad_y = unit(1, "cm"),
      rotation = NULL,
      style = north_arrow_fancy_orienteering
    ) +
    annotation_scale( # add scale bar
      plot_unit = NULL,
      bar_cols = c("black", "white"),
      line_width = 1,
      height = unit(0.15, "cm"),
      pad_x = unit(2, "cm"),
      pad_y = unit(0.15, "cm"),
      text_pad = unit(0.15, "cm"),
      text_cex = 0.7,
      text_face = NULL,
      text_family = "",
      tick_height = 0.6
    ) +
    labs(title = print(paste0("Suitable Percentage for ", species_name))) # add map title 
         
  percent_region
  
  total_area_region+percent_region
}
```


```{r, warning=FALSE}
aquaculture_regions(species_name = "Shrimp", min_temp = 6, max_tem = 12, max_depth = -437, min_depth = -36 )
```

[^hall]:Hall, S. J., Delaporte, A., Phillips, M. J., Beveridge, M. & O’Keefe, M. Blue Frontiers: Managing the Environmental Costs of Aquaculture (The WorldFish Center, Penang, Malaysia, 2011). 

[^gentry]:Gentry, R. R., Froehlich, H. E., Grimm, D., Kareiva, P., Parke, M., Rust, M., Gaines, S. D., & Halpern, B. S. Mapping the global potential for marine aquaculture. *Nature Ecology & Evolution*, 1, 1317-1324 (2017).

[^gebco]:GEBCO Compilation Group (2022) GEBCO_2022 Grid (doi:10.5285/e0f0bb80-ab44-2739-e053-6c86abc0289c).


